# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Set working directory:
``` {r}
echo = FALSE
setwd("~/CourseraHW/RepData_PeerAssessment1")
```
Create a data subdirectory if it does not exist:
``` {r} 
if (!file.exists("data")) {
        dir.create("data")
}
```
Download file and unzip if it isn't already there:
``` {r}
filePath <- "./data/"
fileZipName <- "repdata-data-activity.zip"
fileName <- "activity.csv"
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

ZipFile <- paste0(filePath, fileZipName)        # path to zipped data file
if (!file.exists(ZipFile)) {
        download.file(fileUrl, destfile = ZipFile, method ="curl")
        unzip(ZipFile, exdir=filePath)
        dateDownloaded <- date()
}
```
Read in activity monitoring dataset:
``` {r}
amd <- read.csv(paste0(filePath, fileName))
```

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.
Make a histogram of the total number of steps taken each day.

For each data point, create a column for the day it occurred. Each interval represents a 5-minute total, therefore there are 12 data points per hour and 288 data points per day.
```{r}

```


Calculate and report the mean and median total number of steps taken per day


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
