# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

1. Read in activity monitoring dataset.

Set working directory:
``` {r}
setwd("~/CourseraHW/RepData_PeerAssessment1")
```
Create a data subdirectory if it does not exist:
``` {r} 
if (!file.exists("data")) {
        dir.create("data")
}
```
Download file and unzip if it isn't already there:
``` {r}
filePath <- "./data/"
fileZipName <- "repdata-data-activity.zip"
fileName <- "activity.csv"
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

ZipFile <- paste0(filePath, fileZipName)        # path to zipped data file
if (!file.exists(ZipFile)) {
        download.file(fileUrl, destfile = ZipFile, method ="curl")
        unzip(ZipFile, exdir=filePath)
        dateDownloaded <- date()
}
raw <- read.csv(paste0(filePath, fileName),colClasses = c("integer", "character", "integer"))
```

2. Process/transform the data into a format suitable for analysis
``` {r}
clean <- na.omit(raw)
clean$date <- as.factor(clean$date)
clean$interval <- as.factor(clean$interval)
```

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Make a histogram of the total number of steps taken each day
```{r}
dailysteps <- with(clean, tapply(steps, date, sum))
myhist <- hist(dailysteps, xlab='steps', main='Histogram of Daily Steps')
```

2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
echo = TRUE
cat("Mean Number of Daily Steps = ", mean(dailysteps))
cat("Median Number of Daily Step =", median(dailysteps))
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
intervalsteps <- with(clean, tapply(steps, interval, mean))
times <- as.integer(rownames(intervalsteps))
avgdaily <- cbind(times,intervalsteps)

myplot <- plot(avgdaily, type="l", 
           xlab='5-minute interval', ylab='Number of Steps',
           main='Average Number of Steps')
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
echo = TRUE
cat("The 5-minute Interval with the Maximum Steps is: ", which(avgdaily[,2]==max(avgdaily[,2])))
cat("The Maximum number of steps in this interval =", max(avgdaily[,2]))
```

## Imputing missing values
Note that there are a number of days/intervals where there are missing
values (coded as `NA`). The presence of missing days may introduce
bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
echo = TRUE
cat("Total number of rows with NAs =", (nrow(raw) - nrow(na.omit(raw))))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
fixed <- raw

```


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?
